---
title: "Game Sales Analysis"
author: "Andre Davis"
date: "3/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## install and load the package tidyverse and janitor

library(tidyverse)
library(janitor)
library(ggplot2)
library(gridExtra)
```

*Note*: If you try to Knit this document at this time, you *will* get an error because there is code in this document that has to be edited (by you!) before it will be able to successfully knit!

## Final Project

This exercise has been generated to practice everything you have learned in this course set.

### GitHub Setup

To get started, you'll want to go to GitHub and start a new repository:

- Call this repository `final_project`. 
- Add a short description
- Check the box to "Initialize this repository with a README. 
- Click `Create Repository`

Once the repository has been created, Click on `Clone or download` and copy the "Clone with HTTPS" link provided. You'll use this to clone your repo in RStudio Cloud.

**Note**: If you're stuck on this, these steps were covered in detail in an earlier course: [Version Control](https://leanpub.com/universities/courses/jhu/version-control). Refer to the materials in this course if you're stuck on this part of the project.

### RStudio Cloud Setup

Go to RStudio Coud and create a new project based on Github. As discussed previously, you'll want all your data science projects to be organized from the very beginning. Let's do that now!

First, use `cd` to get yourself into the directory of your GitHub Project.  

Once in the correct directory, use `mkdir` in the terminal to create folders with the following structure:

- data/
  - raw_data/
  - tidy_data/
- code/
  - raw_code/
  - final_code/
- figures/
  - exploratory_figures/
  - explanatory_figures/
- products/
  - writing/

Upload the data file into the tidy_data folder and this .Rmd file into the main project directory.

Once the .Rmd document is in the correct folder, you'll want to **change the author of this document** to your name at the top of the .Rmd document (in the YAML). Save this change before moving to the next step.

**Note**: If you're stuck on this, these steps were covered in detail in an earlier course: [Organizing Data Science Projects](https://leanpub.com/universities/courses/jhu/cbds-organizing). Refer to the materials in this course if you're stuck on this part of the project.

### Pushing to GitHub

You'll want to save changes to your project regularly by pushing them to GitHub. Now that you've got your file structure set up and have added a code file (.Rmd), it's a good time to stage, commit, and push these changes to GitHub. Do so now, and then take a long on GitHub to see the changes on their website!

**Note**: If you're stuck on this, these steps were covered in detail in an earlier course: [Version Control](https://leanpub.com/universities/courses/jhu/version-control). Refer to the materials in this course if you're stuck on this part of the project.


### The data

Video gaming has a long history as far back as the early 1950s but it did not reach mainstream popularity until the 1970s and 1980s when video arcade games and gaming consoles were introduced to the general public. Since the 1980s, video gaming has become a popular form of entertainment and a part of modern culture in most parts of the world. However, the industry also experienced its first major growing pains in early 1980s as publishing houses appeared and competition became intense.

The purpose of this project is to explore the development of the video game industry from 1980-2017, and find market opportunities that industry newcomers may be interested in. Our analysis will focus on publishers, game genres and the popularity of genres by region.

Load the data into a data frame called `gaming`. Make sure you clean the column names so you remove all the caps and spaces in the names. What package do you need to use?


```{r}
gaming <- read_csv("data/tidy_data/Video_Games_Sales_as_at_22_Dec_2016.csv") %>% 
  clean_names()
gaming
```

what are the columns in the data? How many observations are in the data? Write your code in the chunk below.

 
```{r}
colnames(gaming)

```

There might be some missing values in the data. Remove all the missing values from the data and calling the new data frame `gaming_no_na`.

 
```{r}
gaming_no_na <- gaming %>% drop_na()
```


What are the columns in the data. What do you think the column `publisher` contain?

WRITE YOUR ANSWER HERE!! -The columns in this dataset are (name,platform,year_of_release,genre,publisher,na_sales,eu_sales,jp_sales,other_sales,global_sales,critic_score,critic_count,user_score,user_count,developer,rating) The publisher column contains the names of the companines who publish the games.  

The column `publisher` has values such as *Sony Computer Entertainment*, *Sony Computer Entertainment America*, *Sony Computer Entertainment Europe*, *Sony Music Entertainment* and *Sony Online Entertainment*. All those companies are subsidiaries of Sony. Use the appropriate dplyr functon to replace all those values with `Sony`.


```{r}
sony <- c("Sony Computer Entertainment","Sony Computer Entertainment America","Sony Computer Entertainment Europe","Sony Music Entertainment","Sony Online Entertainment")

gaming_no_na <- gaming %>% mutate(publisher = ifelse(publisher %in% sony , "Sony", publisher)) 


```
What are the top 10 publishers by the number of games they produce?


```{r}
gaming_no_na %>% group_by(publisher) %>% 
  tally() %>% 
  top_n(10)   
```
The top 10 publishers based off number of game sales are Activision, Electronic Arts, Konami Digial Entertainment, Namaco Bandi Games, Nintendo, Sega, Sony Computer Entertainment,Take-Two Interactive, THQ, and Ubisoft. 

What are the top 10 publishers by the global sales? The top 10 publishers for global sales are Activision, Electronic Arts, Konmai Digitial Entertainment, Mirosoft Game Studios, Nintento,Sega,Sony,Take-Two Interactive,THQ and Ubisoft

```{r}
top_10<-gaming_no_na %>% 
  group_by(publisher) %>% 
  summarize(total_sales=sum(global_sales)) %>% 
  arrange(-total_sales) %>%
  slice(1:10)   
```

Create a horizontal bar plot using ggplot2 that on the Y axis shows the top 10 companies by overall global sales and on the X axis shows the total global sales. Which company sells the most globally?

```{r}
  
ggplot(top_10,aes(x=reorder(publisher,total_sales),y=total_sales)) + 
  geom_bar(stat="identity",fill="blue")+ labs(x="publisher",y="USD") + coord_flip()
```


Use ggplot2 to create a line graph that shows the toal number of games released each year over time. Do you see any trend?

```{r}
games_per_year<-gaming_no_na %>% 
  group_by(year_of_release) %>% 
  summarize(game_per_year=n()) 

ggplot(games_per_year,aes(x=as.numeric(year_of_release),y=game_per_year))+ labs(x = "year of release",y="games per year") + geom_line()

```


Do the same graph above for the top 10 companies by global sales but group by the company.

```{r}
top_companies<-gaming_no_na %>% 
  group_by(publisher,year_of_release) %>% 
  summarize(total_sales=sum(global_sales))%>%
  filter(publisher %in% top_10$publisher)

ggplot(top_companies,aes(x=as.numeric(year_of_release),y=total_sales,group=publisher,color=publisher))+ labs(x = "year of release",y="total sales")+ geom_line()
```

How about a bar plot that shows the global sales by genres? Which genre sells the most (dollar value)?

```{r}
world_sales<-gaming_no_na %>% 
  group_by(genre) %>% 
  summarize(sales=sum(global_sales))%>%
  drop_na(genre)

ggplot(world_sales, aes(x=reorder(genre,sales), y=sales))+ labs(x ="genre",y="sales USD") + geom_bar(stat="identity", fill="yellow")
```

Do the bar plot above but group them by market (Japan, Europe, North America). Do you see any difference between regions?

```{r}
regional_sales<-gaming_no_na %>% 
  group_by(genre) %>% 
  summarize(sales_na=sum(na_sales),
            sales_eu=sum(eu_sales),
            sales_jp=sum(jp_sales))%>%
  drop_na(genre)

na<-ggplot(regional_sales, aes(x=reorder(genre,sales_na), y=sales_na))+ labs(x = "genre",y="north america sales") + geom_bar(stat="identity", fill="yellow")

eu<-ggplot(regional_sales, aes(x=reorder(genre,sales_eu), y=sales_eu))+ labs(x = "genre",y="europe sales") + geom_bar(stat="identity", fill="yellow")

jp<-ggplot(regional_sales, aes(x=reorder(genre,sales_jp), y=sales_jp))+ labs(x = "genre",y="japan sales") + geom_bar(stat="identity", fill="yellow")

grid.arrange(na,eu,jp,ncol=2)
```

Read the lesson on inferential analysis. We are using linear regression to find what factors determine the sales of a computer game. Use global sales as the dependent variables (Y) and genre, rating, and platform. Make sure to use genre and platform as factors.

```{r}
regression<-lm(global_sales~factor(genre)+rating+factor(platform),data=gaming_no_na)
regression
```


Congratulations! You have completed the project. There are a few final notes:

### Add Markdown Text to .Rmd

Before finalizing your project you'll want be sure there are **comments in your code chunks** and **text outside of your code chunks** to explain what you're doing in each code chunk. These explanations are incredibly helpful for someone who doesn't code or someone unfamiliar to your project.

**Note**: If you're stuck on this, these steps were covered in detail in an earlier course: [Introduction to R](https://leanpub.com/universities/courses/jhu/introduction-to-r). Refer to the R Markdown lesson in this course if you're stuck on this part (or the next part) of the project.


### Knit your R Markdown Document

Last but not least, you'll want to **Knit your .Rmd document into an HTML document**. If you get an error, take a look at what the error says and edit your .Rmd document. Then, try to Knit again! Troubleshooting these error messages will teach you a lot about coding in R.

### A Few Final Checks

A complete project should have:

- Completed code chunks throughout the .Rmd document (your RMarkdown document should Knit without any error)
- README.md text file explaining your project
- Comments in your code chunks
- Answered all questions throughout this exercise.

### Final `push` to GitHub

Now that you've finalized your project, you'll do one final **push to GitHub**. `add`, `commit`, and `push` your work to GitHub. Navigate to your GitHub repository, and answer the final question below! 

**Note**: If you're stuck on this, these steps were covered in detail in an earlier course: [Version Control](https://leanpub.com/universities/courses/jhu/version-control). Refer to the materials in this course if you're stuck on this part of the project.

At the end, submit the link to your github repository to us.

Submit the URL to your `final_project` GitHub repository below.

